Structured Versus Decorated Cospans

### 2021-01-22 cospans.tex
erased %%%
small fixes

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/1/20

@@

Many of the flawed applications of decorated cospans have been fixed using structured cospans \cite[Section 6]{BC}, but not all decorated cospan double categories are equivalent to structured cospan double categories. Here we give sufficient conditions for a decorated cospan double category to be equivalent---and in fact, isomorphic---to a structured cospan double category.

@@

In \cref{DecCospansDoubleCat}, we construct the symmetric monoidal double category $F\lCsp$ and we define maps between decorated cospan double categories. In \cref{EquivDoubleCats}, we briefly review the structured cospans framework and prove that the double categorical versions of decorated cospans and structured cospans are isomorphic under suitable conditions. In \cref{spinoffs}, we establish the isomorphism between structured and decorated cospans at the level of bicategories and categories (via decategorification).  Finally, in \cref{Applications}, we describe applications.

@@

Notice that in the above proof, when we compose two structured cospans, namely two horizontal 1-cells in the double category ${}_L\lCsp(\inta F)$ as in \cref{eq:pushoutissum}, the resulting composite cospan apex ends up being a sum on the second variable due to the definition of the left adjoint functor $L$ which forces the feet of all cospans to involve a fiberwise initial object. Consequently, one could remark that in fact, the assumptions of \cref{thm:equiv} should only state that the corresponding pseudofunctor $F\colon \A\to\SMC$ factors through categories with finite coproducts rather than $\Rex$. We leave the statement as is in order to match the general assumptions of \cref{thm:SC} for the definition of the structured cospan symmetric monoidal double category.

@@

\begin{defn}\label{defn:monoidal_double_functor}
Given symmetric monoidal double categories $\lD$ and $\lE$, a \define{symmetric monoidal double functor} $\lF \maps \lD \to \lE$ is a double functor $\lF$ together with invertible transformations $\lF_\otimes\maps {\otimes \circ (\lF,\lF)} \to \lF \circ \otimes$ and $I_\lE \to \lF \circ I_\lD$ that satisfy the usual coherence axioms for a symmetric monoidal functor.
\end{defn}
\noindent
Explicitly, a symmetric monoidal double functor is a double functor $\lF \maps \lD \to \lE$ such that:
\begin{itemize}
\item{$\lF_0$ and $\lF_1$ are symmetric monoidal functors,}
\item{we have equalities $\lF_0 S_{\lD} = S_{\lE} \lF_1$ and $\lF_0 T_{\lD} = T_{\lE} \lF_1$ of monoidal functors, and}
\item{the following diagrams commute, expressing that $\phi$ is a transformation of double categories:
		\begin{equation}
		\begin{tikzpicture}
			\node (A) at (0,1.5) {\footnotesize{$ (\lF(M_2)\otimes \lF(N_2))\odot (\lF(M_1)\otimes \lF(N_1))$}};
			\node (A') at (0,0) {\footnotesize{$(\lF(M_2) \odot \lF(M_1)) \otimes (\lF(N_2) \odot \lF(N_1))$}};
			\node (B) at (5,1.5) {\footnotesize{$\lF(M_2 \otimes N_2)\odot \lF(M_1 \otimes N_1)$}};
			\node (B') at (5,0) {\footnotesize{$\lF((M_2 \otimes N_2)\odot(M_1 \otimes N_1))$}};
			\node (A'') at (0,-1.5) {\footnotesize{$\lF(M_2 \odot M_1) \otimes \lF(N_2 \odot N_1)$}};
			\node (B'') at (5,-1.5) {\footnotesize{$\lF((M_2 \odot M_1)\otimes(N_2 \odot N_1))$}};
			%
			\path[->,font=\scriptsize]
				(A) edge node[left]{$\chi$} (A')
				(A) edge node[above]{$\lF_\otimes \odot \lF_\otimes$} (B)
				(B) edge node[right]{$\lF_\odot$} (B')
				(A') edge node [left] {$\lF_\odot \otimes \lF_\odot$} (A'')
				(B') edge node [right] {$\lF(\chi)$} (B'')
				(A'') edge node[above]{$\lF_\otimes$} (B'');
		\end{tikzpicture}
		%
		\quad
		%
		\begin{tikzpicture}
			\node (A) at (0,1.5) {\footnotesize{$U_{\lF(a) \otimes \lF(b)}$}};
			\node (A') at (0,0) {\footnotesize{$U_{\lF(a)} \otimes U_{\lF(b)}$}};
			\node (B) at (2,1.5) {\footnotesize{$U_{\lF(a \otimes b)} $}};
			\node (B') at (2,0) {\footnotesize{$\lF(U_{a \otimes b})$}};
			\node (A'') at (0,-1.5) {\footnotesize{$\lF(U_a) \otimes \lF(U_b)$}};
			\node (B'') at (2,-1.5) {\footnotesize{$\lF(U_a \otimes U_b) $}};
			%
			\path[->,font=\scriptsize]
				(A) edge node[left]{$\mu$} (A')
				(A') edge node [left] {$\lF_U \otimes \lF_U$} (A'')
				(B') edge node [right] {$\lF(\mu)$} (B'')
				(A) edge node[above]{$U_{\lF_\otimes}$} (B)
				(B) edge node[right]{$\lF_U$} (B')
				(A'') edge node[above]{$\lF_\otimes$} (B'');
		\end{tikzpicture}
		\end{equation}
}
\end{itemize}
\noindent
%%% Christina: the notation is not streamlined. We occasionally use $F_\otimes$ and $F_\odot$ (which I prefer) and in the definitions above we use various greek letters like $\phi$ (which is definitely overused in our paper). Could this be fixed throughout the document? Similarly for the unit comparison maps.

>>>>

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/1/21

@@

Many of the flawed applications of decorated cospans have been fixed using structured cospans \cite[Section 6]{BC}, but not every decorated cospan double category is equivalent to a structured cospan double category.   Here we give sufficient conditions for a decorated cospan double category to be equivalent---and in fact, isomorphic---to a structured cospan double category.

@@

In \cref{DecCospansDoubleCat}, we construct the symmetric monoidal double category $F\lCsp$ and we define maps between decorated cospan double categories. In \cref{EquivDoubleCats}, we briefly review the structured cospans framework and prove that the double categorical versions of decorated cospans and structured cospans are isomorphic under suitable conditions. In \cref{spinoffs}, we establish the isomorphism between structured and decorated cospans at the level of bicategories and categories (via decategorification).  Finally, in \cref{Applications}, we describe applications, and show that open dynamical system can be described using decorated cospans but not structured cospans.

@@

%Notice that in the above proof, when we compose two structured cospans, namely two horizontal 1-cells in the double category ${}_L\lCsp(\inta F)$ as in \cref{eq:pushoutissum}, the resulting composite cospan apex ends up being a sum on the second variable due to the definition of the left adjoint functor $L$ which forces the feet of all cospans to involve a fiberwise initial object.Thus in fact the assumptions of \cref{thm:equiv} could be weakened to state that the pseudofunctor $F\colon \A\to\SMC$ factors through the 2-category of categories with finite coproducts.

@@

\begin{defn}\label{defn:monoidal_double_functor}
Given symmetric monoidal double categories $\lD$ and $\lE$, a \define{symmetric monoidal double functor} $\lF \maps \lD \to \lE$ is a double functor $\lF$ together with invertible transformations $\lF_\otimes\maps {\otimes \circ (\lF,\lF)} \to \lF \circ \otimes$ and $I_\lE \to \lF \circ I_\lD$ that satisfy the usual coherence axioms for a symmetric monoidal functor.
\end{defn}
\noindent
Explicitly, a symmetric monoidal double functor is a double functor $\lF \maps \lD \to \lE$ such that:
\begin{itemize}
\item{$\lF_0$ and $\lF_1$ are symmetric monoidal functors,}
\item{we have equalities $\lF_0 S_{\lD} = S_{\lE} \lF_1$ and $\lF_0 T_{\lD} = T_{\lE} \lF_1$ of monoidal functors, and}
\item{the following diagrams commute, expressing that $\phi$ is a transformation of double categories:
		\begin{equation}
		\begin{tikzpicture}
			\node (A) at (0,1.5) {\footnotesize{$ (\lF(M_2)\otimes \lF(N_2))\odot (\lF(M_1)\otimes \lF(N_1))$}};
			\node (A') at (0,0) {\footnotesize{$(\lF(M_2) \odot \lF(M_1)) \otimes (\lF(N_2) \odot \lF(N_1))$}};
			\node (B) at (5,1.5) {\footnotesize{$\lF(M_2 \otimes N_2)\odot \lF(M_1 \otimes N_1)$}};
			\node (B') at (5,0) {\footnotesize{$\lF((M_2 \otimes N_2)\odot(M_1 \otimes N_1))$}};
			\node (A'') at (0,-1.5) {\footnotesize{$\lF(M_2 \odot M_1) \otimes \lF(N_2 \odot N_1)$}};
			\node (B'') at (5,-1.5) {\footnotesize{$\lF((M_2 \odot M_1)\otimes(N_2 \odot N_1))$}};
			%
			\path[->,font=\scriptsize]
				(A) edge node[left]{$\chi$} (A')
				(A) edge node[above]{$\lF_\otimes \odot \lF_\otimes$} (B)
				(B) edge node[right]{$\lF_\odot$} (B')
				(A') edge node [left] {$\lF_\odot \otimes \lF_\odot$} (A'')
				(B') edge node [right] {$\lF(\chi)$} (B'')
				(A'') edge node[above]{$\lF_\otimes$} (B'');
		\end{tikzpicture}
		%
		\quad
		%
		\begin{tikzpicture}
			\node (A) at (0,1.5) {\footnotesize{$U_{\lF(a) \otimes \lF(b)}$}};
			\node (A') at (0,0) {\footnotesize{$U_{\lF(a)} \otimes U_{\lF(b)}$}};
			\node (B) at (2,1.5) {\footnotesize{$U_{\lF(a \otimes b)} $}};
			\node (B') at (2,0) {\footnotesize{$\lF(U_{a \otimes b})$}};
			\node (A'') at (0,-1.5) {\footnotesize{$\lF(U_a) \otimes \lF(U_b)$}};
			\node (B'') at (2,-1.5) {\footnotesize{$\lF(U_a \otimes U_b) $}};
			%
			\path[->,font=\scriptsize]
				(A) edge node[left]{$\mu$} (A')
				(A') edge node [left] {$\lF_U \otimes \lF_U$} (A'')
				(B') edge node [right] {$\lF(\mu)$} (B'')
				(A) edge node[above]{$U_{\lF_\otimes}$} (B)
				(B) edge node[right]{$\lF_U$} (B')
				(A'') edge node[above]{$\lF_\otimes$} (B'');
		\end{tikzpicture}
		\end{equation}
}
\end{itemize}
